Theory about Normal Basis on $F(2^k)$

### 2017-04-02 dissertation/submission/intro.tex
dissertation: SAGA ends. upload to proQuest

Boolean functions with $k$-bit inputs and $k$-bit outputs 
$f: \B^k \rightarrow \B^k,\B = \{0, 1\}$ can be construed as functions
$f: \Fkk \rightarrow \Fkk$. It is well known that over the finite
field ($\Fq$) of  $q$ elements, every function $f: \Fq
\rightarrow \Fq$ is a polynomial function \cite{ff:1997}. Moreover,
there exists a unique canonical polynomial $\Func$ that describes $f$.
This implies that one can derive a canonical, polynomial
  abstraction of the function as $Z = \Func(A)$ where $Z, A$ are
word-level symbols representing $k$-bit operands. The concept also
generalizes to functions with different input/output bit-vector sizes,
{\it i.e.} functions of the type $f: \B^n \rightarrow \B^m$, modeled as
polynomials over $f:{\mathbb{F}}_{2^k} \rightarrow
{\mathbb{F}}_{2^k}$, where $k=LCM(n,m)$ \cite{ff:1997}.  
{\it This implies that the FSM's transition relations can be
represented as polynomial functions (ideals) in $\Fkk$, and values of
state variables can be represented as solutions to these polynomials}
(variety of the ideal). Subsequently, the ideal-variety correspondences
in algebraic geometry can be applied to implement symbolic reasoning
about state reachability. 
% Moreover, as ${\mathbb{F}}_2 \subset \Fkk$,
% our model provides a single, unified and bit-precise representation
% for both bit-level (${\mathbb{F}}_2$) and word-level ($\Fkk$)
% constraints.

@@

\subsection{Application to Sequential Galois Field Arithmetic Circuit Verification}
Sequential Galois field (GF) arithmetic circuits can be modeled as a special type of 
FSM, where the preloaded operands are mapped to initial states, and pseudo outputs
after $k$ clock-cycles are recognized as the final reached state after $k$ transitions.
Therefore, the word-level FSM traversal algorithm can be applied to 
verify the correctness of final reached state, {\it i.e.} the functional correctness of
a sequential arithmetic circuit.

>>>>

Boolean functions with $k$-bit inputs and $k$-bit outputs 
$f: \B^k \rightarrow \B^k,\B = \{0, 1\}$ can be construed as functions
$f: \Fkk \rightarrow \Fkk$. It is well known that over the finite
field ($\Fq$) of  $q$ elements, every function $f: \Fq
\rightarrow \Fq$ is a polynomial function \cite{ff:1997}. Moreover,
there exists a unique canonical polynomial $\Func$ that describes $f$.
This implies that one can derive a canonical, polynomial
  abstraction of the function as $Z = \Func(A)$ where $Z, A$ are
word-level symbols representing $k$-bit operands. The concept also
generalizes to functions with different input/output bit-vector sizes,
i.e., functions of the type $f: \B^n \rightarrow \B^m$, modeled as
polynomials over $f:{\mathbb{F}}_{2^k} \rightarrow
{\mathbb{F}}_{2^k}$, where $k=LCM(n,m)$ \cite{ff:1997}.  
{\it This implies that the FSM's transition relations can be
represented as polynomial functions (ideals) in $\Fkk$, and values of
state variables can be represented as solutions to these polynomials}
(variety of the ideal). Subsequently, the ideal-variety correspondences
in algebraic geometry can be applied to implement symbolic reasoning
about state reachability. 
% Moreover, as ${\mathbb{F}}_2 \subset \Fkk$,
% our model provides a single, unified and bit-precise representation
% for both bit-level (${\mathbb{F}}_2$) and word-level ($\Fkk$)
% constraints.

@@

\subsection{Application to Sequential Galois Field Arithmetic Circuit Verification}
Sequential Galois field (GF) arithmetic circuits can be modeled as a special type of 
FSM, where the preloaded operands are mapped to initial states, and pseudo outputs
after $k$ clock-cycles are recognized as the final reached state after $k$ transitions.
Therefore, the word-level FSM traversal algorithm can be applied to 
verify the correctness of final reached state, i.e., the functional correctness of
a sequential arithmetic circuit.

### 2017-04-02 dissertation/submission/prelim_GF.tex
dissertation: SAGA ends. upload to proQuest

\begin{Definition}
A {\bf field} $\mathbb{F}$ is a commutative ring with unity, where every
nonzero element in $\mathbb{F}$ has a multiplicative inverse; {\it i.e.} $\forall
a \in \mathbb{F} - \{0\}$, $\exists \hat{a} \in \mathbb{F}$ such that $ a \cdot
\hat{a} = 1$.
\end{Definition}

@@

\begin{Definition} 
A {\bf Galois field}, denote $\Fq$, is a field with a finite
number of elements, $q$. The number of elements $q$ of the field is
a power of a prime integer, {\it i.e.} $q = p^k$, where $p$ is a prime
integer, and $k \geq 1$. Thus a Galois field can also be denoted as 
$\F_{p^{k}}$.
\end{Definition}

@@

\begin{Theorem}
$\F_{2^n}\subset\F_{2^m}$ iff $n \mid m$, {\it i.e.} if $n$ divides $m$.
\end{Theorem}

@@

The final remainder is $1$, {\it i.e.} the product equals 1 in StdB.

>>>>

\begin{Definition}
A {\bf field} $\mathbb{F}$ is a commutative ring with unity, where every
nonzero element in $\mathbb{F}$ has a multiplicative inverse; i.e., $\forall
a \in \mathbb{F} - \{0\}$, $\exists \hat{a} \in \mathbb{F}$ such that $ a \cdot
\hat{a} = 1$.
\end{Definition}

@@

\begin{Definition} 
A {\bf Galois field}, denote $\Fq$, is a field with a finite
number of elements, $q$. The number of elements $q$ of the field is
a power of a prime integer, i.e., $q = p^k$, where $p$ is a prime
integer, and $k \geq 1$. Thus a Galois field can also be denoted as 
$\F_{p^{k}}$.
\end{Definition}

@@

\begin{Theorem}
$\F_{2^n}\subset\F_{2^m}$ iff $n \mid m$, {\it i.e.}, if $n$ divides $m$.
\end{Theorem}

@@

The final remainder is $1$, i.e., the product equals 1 in StdB.

### 2017-04-02 dissertation/submission/prelim_basis.tex
dissertation: after submission_final revision
dissertation: SAGA ends. upload to proQuest

A {\bf lexicographic ordering} (LEX) is a total-ordering $>$ such that 
variables in the terms are lexicographically ordered, {\it i.e.} simply based on 
when the variables appear in the ordering. It is also a well-order, where the least element exists.
Higher variable-degrees take 
precedence over lower degrees for equivalent variables ({\it e.g.} $a^3 > a^2$ due to $a \cdot a \cdot a > a \cdot a \cdot 1$).
\begin{Definition}
{\bf Lexicographic order:} Let $x_1 > x_2 > \dots > x_d$
lexicographically. Also let $\alpha = (\alpha_1, \dots, \alpha_d);
~\beta = (\beta_1, \dots, \beta_d) \in \mathbb{Z}^d_{\geq 0}$. Then we
have: 
\begin{equation}
x^{\alpha} > x^{\beta} \iff 
\begin{cases}
& \text{Starting  from the  left, the first co-ordinates of $\alpha_i, \beta_i$} \\
& \text{that are different satisfy $\alpha_i > \beta_i$}

@@

Let $\mathbb{F}$ be any field and let $\mathbf{a}=(a_{1},\dots,a_{d}) \in \mathbb{F}^d$ be a point, and $f \in
\mathbb{F}[x_1,\dots, x_d]$ be a polynomial. We say that $f$ {\it vanishes} on $\mathbf{a}$ 
if $f(\mathbf{a}) = 0$, {\it i.e.} $\mathbf{a}$ is in the variety of $f$.

@@

\begin{Definition}\label{def:minigb}
A {\bf minimal Gr\"obner basis} for a polynomial ideal $I$ is a \Grobner basis $G$ for $I$ such that
	\begin{itemize}
		\item $lc(g_{i})=1,\forall g_{i}\in G$
		\item $\forall g_{i} \in G$,  $lt(g_{i}) \notin \langle lt(G-\{g_{i}\})\rangle$
	\end{itemize}
\end{Definition}
A {\bf minimal} \Grobner basis is a \Grobner basis such that all polynomials
have a coefficient of $1$ and no leading term of any element in $G$ divides 
another in $G$.
Given a \Grobner basis $G$, a minimal \Grobner basis can be
computed as follows:
\begin{enumerate}
\item Minimize every $g_i \in G$, i.e., $g_i=g_i/lc(g_i)$
\item For $g_i, g_j \in G$ where $i\neq j$, remove $g_i$ from $G$ if $lt(g_i)\mid lt(g_j)$, 
i.e., remove every polynomial in $G$ whose leading term is divisible by the leading term 
of some other polynomial in $G$.
\end{enumerate}

@@

{\bf Weak Nullstellensatz} offers a way to evaluate whether or not the 
system of multivariate polynomial equations (ideal $J$) has common solutions 
in ${\overline {\F}}^d$. For this purpose, we only need to check if 
the ideal is generated by the unit element, {\it i.e.} $1\in J$. 
This approach can be used to evaluate the feasibility of constraints in 
verification problems.
%%%%%%%%%%%%%%%%%%%%%%%%%strong Nullstellensatz%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
An interesting result is one of {\bf Strong 
Nullstellensatz}.
The strong
Nullstellensatz establishes the correspondence between radical ideals
and varieties.

@@

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/abs_setup.pdf}
\caption{Galois field arithmetic circuit model for abstraction.}
\label{fig:abs_setup}}
\end{figure}

>>>>

A {\bf lexicographic ordering} (LEX) is a total-ordering $>$ such that 
variables in the terms are lexicographically ordered, i.e., simply based on 
when the variables appear in the ordering. It is also a well-order, where the least element exists.
Higher variable-degrees take 
precedence over lower degrees for equivalent variables (e.g., $a^3 > a^2$ due to $a \cdot a \cdot a > a \cdot a \cdot 1$).
\begin{Definition}
{\bf Lexicographic order:} Let $x_1 > x_2 > \dots > x_d$
lexicographically. Also let $\alpha = (\alpha_1, \dots, \alpha_d);
~\beta = (\beta_1, \dots, \beta_d) \in \mathbb{Z}^d_{\geq 0}$. Then we
have: 
\begin{equation}
x^{\alpha} > x^{\beta} \iff 
\begin{cases}
& \text{Starting  from the  left, the first co-ordinates of $\alpha_i, \beta_i$} \\
& \text{that are different satisfy $\alpha_i > \beta_i$}

@@

Let $\mathbb{F}$ be any field and let $\mathbf{a}=(a_{1},\dots,a_{d}) \in \mathbb{F}^d$ be a point, and $f \in
\mathbb{F}[x_1,\dots, x_d]$ be a polynomial. We say that $f$ {\it vanishes} on $\mathbf{a}$ 
if $f(\mathbf{a}) = 0$, i.e., $\mathbf{a}$ is in the variety of $f$.

@@

\begin{Definition}\label{def:minigb}
A {\bf minimal Gr\"obner basis} for a polynomial ideal $I$ is a \Grobner basis $G$ for $I$ such that
	\begin{itemize}
		\item $lc(g_{i})=1,\forall g_{i}\in G$
		\item $\forall g_{i} \in G$,  $lt(g_{i}) \notin \langle lt(G-\{g_{i}\})\rangle$
	\end{itemize}
\end{Definition}
A {\bf minimal} \Grobner basis is a \Grobner basis such that all polynomials
have a coefficient of $1$ and no leading term of any element in $G$ divides 
another in $G$.
Given a \Grobner basis $G$, a minimal \Grobner basis can be
computed as follows:
\begin{enumerate}
\item Minimize every $g_i \in G$, i.e., $g_i=g_i/lc(g_i)$.
\item For $g_i, g_j \in G$ where $i\neq j$, remove $g_i$ from $G$ if $lt(g_i)\mid lt(g_j)$, 
i.e., remove every polynomial in $G$ whose leading term is divisible by the leading term 
of some other polynomial in $G$.
\end{enumerate}

@@

{\bf Weak Nullstellensatz} offers a way to evaluate whether or not the 
system of multivariate polynomial equations (ideal $J$) has common solutions 
in ${\overline {\F}}^d$. For this purpose, we only need to check if 
the ideal is generated by the unit element, i.e., $1\in J$. 
This approach can be used to evaluate the feasibility of constraints in 
verification problems.
%%%%%%%%%%%%%%%%%%%%%%%%%strong Nullstellensatz%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
An interesting result is one of {\bf Strong 
Nullstellensatz}.
The strong
Nullstellensatz establishes the correspondence between radical ideals
and varieties.

@@

\begin{figure}[tbp]
\centering{
\includegraphics[width=\textwidth]{newfig/abs_setup.pdf}
\caption{Galois field arithmetic circuit model for abstraction.}
\label{fig:abs_setup}}
\end{figure}
